---
title: "TFH-to-TR1 multimodal preprocessing"
author: "Enric Canut Zimmermann"
output: html_document
---

# Purpose

This notebook contains the stable preprocessing steps used to build QC-filtered multimodal Seurat objects from TF-specific single-cell CUT&Tag and matched gene-expression data. Exploratory QC plots and abandoned analysis branches from the development notebook have been removed.

# 1. Setup

```{r setup, warning=FALSE, message=FALSE}
library(Signac)
library(Seurat)
library(Matrix)
library(GenomicRanges)
library(EnsDb.Mmusculus.v79)
library(BSgenome.Mmusculus.UCSC.mm10)
library(dplyr)
library(ggplot2)
library(qs)

genome_name <- "mm10"
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevels(annotation) <- paste0("chr", seqlevels(annotation))
genome(annotation) <- genome_name

samples <- c("Bcl6", "Irf4", "Maf")
min_cells_assay <- 1
min_features_assay <- 0

project_data <- file.path("data", "sp20_21")
ct_root <- file.path(project_data, "relaxed", "scCT")
rna_root <- file.path(project_data, "relaxed", "rna")

sample_dirs <- setNames(
  file.path(ct_root, samples),
  samples
)

rna_dirs <- setNames(
  file.path(rna_root, samples),
  samples
)

dir.create(rna_root, recursive = TRUE, showWarnings = FALSE)
dir.create(file.path("figures", "qc"), recursive = TRUE, showWarnings = FALSE)
```

# 2. Load TF-specific peak calls and cell metadata

```{r load-cuttag-inputs}
peak_ranges <- list()
metadata <- list()
fragments <- list()

for (tf in samples) {
  peak_file <- file.path(sample_dirs[[tf]], "macs2_permissive_peaks.narrowPeak")
  metadata_file <- file.path(sample_dirs[[tf]], "picked_cellsmetadata_f.csv")
  fragment_file <- file.path(sample_dirs[[tf]], "fragments.tsv.gz")

  stopifnot(
    file.exists(peak_file),
    file.exists(metadata_file),
    file.exists(fragment_file)
  )

  peak_df <- read.table(peak_file)[, 1:3]
  colnames(peak_df) <- c("chr", "start", "end")
  peak_ranges[[tf]] <- reduce(makeGRangesFromDataFrame(peak_df))

  meta <- read.csv(metadata_file, stringsAsFactors = FALSE)
  stopifnot("barcode" %in% colnames(meta))
  rownames(meta) <- meta$barcode
  metadata[[tf]] <- meta

  fragments[[tf]] <- CreateFragmentObject(
    path = fragment_file,
    cells = meta$barcode
  )
}
```

# 3. Initial permissive CUT&Tag quantification

The development workflow used a pooled permissive reference at this preprocessing stage. A final shared consensus reference is rebuilt after RNA-based state annotation and QC in the thesis-analysis notebook.

```{r initial-cuttag-quantification}
initial_union_peaks <- reduce(unlist(GRangesList(peak_ranges)))

cuttag_objects <- list()

for (tf in samples) {
  counts <- FeatureMatrix(
    fragments = fragments[[tf]],
    features = initial_union_peaks,
    cells = metadata[[tf]]$barcode,
    process_n = 20000
  )

  chrom_assay <- CreateChromatinAssay(
    counts = counts,
    fragments = fragments[[tf]],
    genome = genome_name,
    min.cells = min_cells_assay,
    min.features = min_features_assay,
    annotation = annotation
  )

  cuttag_objects[[tf]] <- CreateSeuratObject(
    counts = chrom_assay,
    assay = "peaks",
    meta.data = metadata[[tf]],
    project = tf
  )

  cuttag_objects[[tf]]$sample <- tf
}
```

# 4. Map CUT&Tag barcodes to matched gene-expression barcodes

```{r barcode-translation}
atac_whitelist_path <- file.path(project_data, "scCT", "737K-arc-v1.txt.gz")
gex_whitelist_path <- file.path(project_data, "rna", "737K-arc-v1.txt.gz")

atac_whitelist <- readLines(gzfile(atac_whitelist_path))
gex_whitelist <- readLines(gzfile(gex_whitelist_path))
stopifnot(length(atac_whitelist) == length(gex_whitelist))

barcode_translation <- data.frame(
  ATAC_barcode = paste0(atac_whitelist, "-1"),
  GEX_barcode = paste0(gex_whitelist, "-1"),
  stringsAsFactors = FALSE
)
```

# 5. Build matched multimodal Seurat objects

```{r build-joint-objects}
for (tf in samples) {
  rna_raw <- Read10X(data.dir = rna_dirs[[tf]])
  ct_obj <- cuttag_objects[[tf]]

  peak_ranges_std <- keepStandardChromosomes(
    granges(ct_obj[["peaks"]]),
    pruning.mode = "coarse"
  )
  ct_obj <- subset(
    ct_obj,
    features = GRangesToString(peak_ranges_std),
    assay = "peaks"
  )

  old_cells <- Cells(ct_obj[["peaks"]])
  matches <- match(old_cells, barcode_translation$ATAC_barcode)

  if (anyNA(matches)) {
    warning(sum(is.na(matches)), " CUT&Tag barcodes did not match the ARC whitelist for ", tf)
  }

  keep <- !is.na(matches)
  ct_obj <- subset(ct_obj, cells = old_cells[keep])
  new_cells <- barcode_translation$GEX_barcode[matches[keep]]
  ct_assay <- RenameCells(ct_obj[["peaks"]], new.names = new_cells)

  master_cells <- intersect(Cells(ct_assay), colnames(rna_raw))
  message(tf, ": ", length(master_cells), " matched multimodal barcodes")

  rna_sub <- rna_raw[, master_cells, drop = FALSE]
  joint <- CreateSeuratObject(counts = rna_sub, project = tf)

  ct_counts <- GetAssayData(ct_assay, slot = "counts")[, master_cells, drop = FALSE]
  joint[["peaks"]] <- CreateChromatinAssay(
    counts = ct_counts,
    ranges = granges(ct_assay),
    fragments = Fragments(ct_assay)
  )

  joint$sampletype <- paste0("multi.", tf)
  qsave(joint, file.path(rna_root, paste0("joint_", tf, "_relaxed.qs")))
}
```

# 6. Add QC metrics

```{r qc-metrics}
for (tf in samples) {
  input_file <- file.path(rna_root, paste0("joint_", tf, "_relaxed.qs"))
  obj <- qread(input_file)

  DefaultAssay(obj) <- "RNA"
  obj[["percent.mt"]] <- PercentageFeatureSet(obj, pattern = "^mt-")
  obj[["percent.rb"]] <- PercentageFeatureSet(obj, pattern = "^Rp")

  qsave(obj, file.path(rna_root, paste0("joint_", tf, "_QC_added.qs")))
}
```

# 7. Apply final RNA QC thresholds

```{r final-qc-filtering}
qc_params <- list(
  Bcl6 = list(min_rna = 450, max_rna = 10000, max_features = 7500, max_mt = 3),
  Irf4 = list(min_rna = 450, max_rna = 10000, max_features = 7500, max_mt = 3),
  Maf  = list(min_rna = 300, max_rna = 10000, max_features = 7500, max_mt = 3)
)

qc_summary <- list()

for (tf in samples) {
  obj <- qread(file.path(rna_root, paste0("joint_", tf, "_QC_added.qs")))
  pars <- qc_params[[tf]]

  filtered <- subset(
    obj,
    subset =
      nCount_RNA > pars$min_rna &
      nCount_RNA < pars$max_rna &
      nFeature_RNA < pars$max_features &
      percent.mt < pars$max_mt
  )

  qc_summary[[tf]] <- data.frame(
    TF = tf,
    Cells_before = ncol(obj),
    Cells_after = ncol(filtered)
  )

  qsave(
    filtered,
    file.path(rna_root, paste0("joint_", tf, "_filt_100726.qs"))
  )
}

qc_summary <- bind_rows(qc_summary)
write.csv(qc_summary, file.path("results", "tables", "qc_cell_counts.csv"), row.names = FALSE)
```

# 8. QC plots used for the supplementary material

```{r qc-plots}
filtered_objects <- setNames(
  lapply(samples, function(tf) {
    qread(file.path(rna_root, paste0("joint_", tf, "_filt_100726.qs")))
  }),
  samples
)

for (tf in samples) {
  obj <- filtered_objects[[tf]]

  p1 <- VlnPlot(
    obj,
    features = c(
      "nCount_peaks", "nFeature_peaks",
      "nCount_RNA", "nFeature_RNA",
      "percent.mt", "percent.rb"
    ),
    ncol = 3,
    pt.size = 0
  ) + ggtitle(paste(tf, "post-filter distributions"))

  p2 <- FeatureScatter(
    obj,
    feature1 = "nCount_RNA",
    feature2 = "nFeature_RNA",
    pt.size = 0.5
  ) + ggtitle(paste(tf, "RNA QC"))

  ggsave(
    file.path("figures", "qc", paste0("QC_", tf, "_violin.png")),
    p1,
    width = 10,
    height = 7,
    dpi = 300
  )
  ggsave(
    file.path("figures", "qc", paste0("QC_", tf, "_rna_scatter.png")),
    p2,
    width = 6,
    height = 5,
    dpi = 300
  )
}
```

# 9. Reproducibility

```{r session-info}
writeLines(capture.output(sessionInfo()), "sessionInfo_preprocessing.txt")
```
