---
title: "TFH-to-TR1 scCUT&Tag multiome analysis"
author: "Enric Canut Zimmermann"
output: html_document
---

# Purpose

This notebook contains the analyses that directly support the final Master's thesis. It starts from QC-filtered multimodal Seurat objects and excludes exploratory analyses that were not retained in the final interpretation (promoter-level aggregation, motif enrichment, and cross-TF co-binding).

# 1. Setup

```{r setup, warning=FALSE, message=FALSE}
library(Seurat)
library(Signac)
library(Matrix)
library(GenomicRanges)
library(EnsDb.Mmusculus.v79)
library(BSgenome.Mmusculus.UCSC.mm10)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(ggplot2)
library(patchwork)
library(scales)
library(qs)

set.seed(123)

genome_name <- "mm10"
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevels(annotation) <- paste0("chr", seqlevels(annotation))
genome(annotation) <- genome_name

state_order <- c("TFH", "TR1.1", "TR1.2", "Foxp3+TR1.2")
state_order_with_undefined <- c(state_order, "Undefined")

state_cols <- c(
  "TFH" = "#F8766D",
  "TR1.1" = "#7CAE00",
  "TR1.2" = "#00BFC4",
  "Foxp3+TR1.2" = "#C77CFF",
  "Undefined" = "grey54"
)

# Adjust these paths locally if required.
data_dir <- file.path("data", "sp20_21", "relaxed", "rna")
reference_atlas_path <- file.path("data", "ctls_only_obj_260126.qs")

input_files <- c(
  Bcl6 = file.path(data_dir, "joint_Bcl6_filt_100726.qs"),
  Irf4 = file.path(data_dir, "joint_Irf4_filt_100726.qs"),
  Maf  = file.path(data_dir, "joint_Maf_filt_100726.qs")
)

dir.create("figures", recursive = TRUE, showWarnings = FALSE)
dir.create(file.path("results", "tables"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path("results", "objects"), recursive = TRUE, showWarnings = FALSE)
```

# 2. Load QC-filtered multimodal objects

```{r load-objects}
tf_objects <- lapply(input_files, qread)
stopifnot(identical(names(tf_objects), c("Bcl6", "Irf4", "Maf")))

atlas <- qread(reference_atlas_path)
atlas@meta.data <- atlas@meta.data %>%
  mutate(
    pops = case_when(
      celltype == "TFH" ~ "TFH",
      celltype == "TR1-like" ~ "TR1.1",
      celltype == "TR1" ~ "TR1.2",
      celltype == "Foxp3+TR1" ~ "Foxp3+TR1.2",
      TRUE ~ NA_character_
    )
  )
DefaultAssay(atlas) <- "integrated"
```

# 3. RNA normalization and reference-based cell-state annotation

```{r rna-annotation}
for (tf in names(tf_objects)) {
  current_obj <- tf_objects[[tf]]
  DefaultAssay(current_obj) <- "RNA"

  current_obj <- NormalizeData(current_obj, verbose = FALSE)
  current_obj <- FindVariableFeatures(current_obj, verbose = FALSE)
  current_obj <- ScaleData(current_obj, verbose = FALSE)
  current_obj <- RunPCA(current_obj, verbose = FALSE)

  anchors <- FindTransferAnchors(
    reference = atlas,
    query = current_obj,
    dims = 1:30,
    reference.reduction = "pca",
    verbose = FALSE
  )

  predictions <- TransferData(
    anchorset = anchors,
    refdata = atlas$pops,
    dims = 1:30,
    verbose = FALSE
  )

  current_obj <- AddMetaData(current_obj, metadata = predictions)
  current_obj$celltype <- as.character(current_obj$predicted.id)
  current_obj$celltype[current_obj$prediction.score.max < 0.5] <- "Undefined"
  current_obj$celltype <- factor(
    current_obj$celltype,
    levels = state_order_with_undefined
  )

  current_obj <- RunUMAP(
    current_obj,
    reduction = "pca",
    dims = 1:30,
    reduction.name = "umap_rna",
    verbose = FALSE
  )

  tf_objects[[tf]] <- current_obj
}
```

# 4. State-specific gene-signature scoring

```{r module-scores}
gene_signatures <- list(
  TFH = c(
    "Art2b", "Ascl2", "Bcl6", "Cdc25b", "Cxcr5", "Cxxc5", "Id3", "Il4",
    "Matk", "Nav2", "Nfia", "Ntrk3", "Pdcd1", "Pde2a", "Ppp1r14b", "Prag1",
    "Rgcc", "Tox2"
  ),
  TR1.1 = c(
    "Ahnak", "Ccr7", "Cd200", "Cd226", "Dapl1", "Edn3", "Emp3", "Gm2682",
    "Klf2", "Nrn1", "Rasa3", "Rasgrp2", "S100a6", "S1pr1", "Vim", "Zbtb32"
  ),
  TR1.2 = c(
    "Adam19", "Arl5a", "Coro2a", "Cxcr6", "Gimap7", "Il10", "Nkg7", "Ppp3ca",
    "Prdm1", "Rgs1", "S100a6"
  ),
  `Foxp3+TR1.2` = c(
    "Ccl5", "Ccr2", "Csf1", "Fgl2", "Foxp3", "Gzmb", "Hip1", "Nkg7", "S100a4"
  )
)

for (tf in names(tf_objects)) {
  current_obj <- tf_objects[[tf]]
  DefaultAssay(current_obj) <- "RNA"

  current_obj <- AddModuleScore(
    current_obj,
    features = gene_signatures,
    name = "SigScore_"
  )

  current_obj$TFH_Score <- current_obj$SigScore_1
  current_obj$TR1.1_Score <- current_obj$SigScore_2
  current_obj$TR1.2_Score <- current_obj$SigScore_3
  current_obj$Foxp3_TR1.2_Score <- current_obj$SigScore_4

  tf_objects[[tf]] <- current_obj
}
```

# 5. CUT&Tag TF-IDF normalization and LSI

```{r cuttag-lsi}
lsi_params <- list(
  Bcl6 = list(n_svd = 9, umap_dims = 2:9),
  Irf4 = list(n_svd = 9, umap_dims = 2:9),
  Maf  = list(n_svd = 3, umap_dims = 2:3)
)

for (tf in names(tf_objects)) {
  current_obj <- tf_objects[[tf]]
  pars <- lsi_params[[tf]]

  DefaultAssay(current_obj) <- "peaks"
  current_obj <- RunTFIDF(current_obj, verbose = FALSE)
  current_obj <- FindTopFeatures(
    current_obj,
    min.cutoff = "q5",
    verbose = FALSE
  )
  current_obj <- RunSVD(current_obj, n = pars$n_svd, verbose = FALSE)
  current_obj <- RunUMAP(
    current_obj,
    reduction = "lsi",
    dims = pars$umap_dims,
    reduction.name = "umap_cuttag",
    verbose = FALSE
  )

  tf_objects[[tf]] <- current_obj
}
```

# 6. Build the shared consensus peak reference

```{r consensus-peaks}
original_peak_ranges <- lapply(
  tf_objects,
  function(x) granges(x[["peaks"]])
)

consensus_peaks_unfiltered <- reduce(
  unlist(GRangesList(original_peak_ranges))
)

peak_width_summary <- summary(width(consensus_peaks_unfiltered))
peak_width_quantiles <- quantile(
  width(consensus_peaks_unfiltered),
  probs = c(0.90, 0.95, 0.99, 0.999)
)

n_over_2kb <- sum(width(consensus_peaks_unfiltered) > 2000)
pct_over_2kb <- 100 * n_over_2kb / length(consensus_peaks_unfiltered)

consensus_peaks <- consensus_peaks_unfiltered[
  width(consensus_peaks_unfiltered) <= 2000
]

common_counts <- lapply(names(tf_objects), function(tf) {
  FeatureMatrix(
    fragments = Fragments(tf_objects[[tf]][["peaks"]]),
    features = consensus_peaks,
    cells = colnames(tf_objects[[tf]])
  )
})
names(common_counts) <- names(tf_objects)

for (tf in names(tf_objects)) {
  tf_objects[[tf]][["peaks_common"]] <- CreateChromatinAssay(
    counts = common_counts[[tf]],
    fragments = Fragments(tf_objects[[tf]][["peaks"]]),
    genome = genome_name,
    annotation = annotation,
    min.cells = 0,
    min.features = 0
  )
}

# Retain a region when it is detected in at least one cell in at least one TF dataset.
keep_peak <- Reduce(
  `|`,
  lapply(tf_objects, function(x) {
    mat <- GetAssayData(x, assay = "peaks_common", slot = "counts")
    Matrix::rowSums(mat > 0) >= 1
  })
)

consensus_peaks_final <- consensus_peaks[keep_peak]
consensus_names <- GRangesToString(consensus_peaks_final)

for (tf in names(tf_objects)) {
  tf_objects[[tf]][["peaks_common"]] <- subset(
    tf_objects[[tf]][["peaks_common"]],
    features = consensus_names
  )
}

stopifnot(length(unique(vapply(
  tf_objects,
  function(x) nrow(x[["peaks_common"]]),
  numeric(1)
))) == 1)
```

# 7. Pseudobulk coverage characteristics

```{r pseudobulk-qc}
pseudobulk_qc <- imap_dfr(tf_objects, function(current_obj, tf_name) {
  current_obj <- subset(current_obj, subset = celltype != "Undefined")
  DefaultAssay(current_obj) <- "peaks_common"

  pb_counts <- AggregateExpression(
    current_obj,
    assays = "peaks_common",
    group.by = "celltype",
    slot = "counts",
    return.seurat = FALSE
  )$peaks_common

  n_cells <- table(current_obj$celltype)
  total_peak_counts <- colSums(pb_counts)
  n_detected_peaks <- colSums(pb_counts > 0)

  pb_cpm <- sweep(pb_counts, 2, total_peak_counts, FUN = "/") * 1e6
  n_peaks_cpm10 <- colSums(pb_cpm > 10)

  tibble(
    TF = tf_name,
    State = colnames(pb_counts),
    N_cells = as.numeric(n_cells[colnames(pb_counts)]),
    Total_peak_counts = as.numeric(total_peak_counts),
    Counts_per_cell = as.numeric(total_peak_counts) /
      as.numeric(n_cells[colnames(pb_counts)]),
    N_detected_peaks = as.numeric(n_detected_peaks),
    `N_peaks_CPM>10` = as.numeric(n_peaks_cpm10)
  )
}) %>%
  mutate(State = factor(State, levels = state_order)) %>%
  arrange(factor(TF, levels = c("Bcl6", "Irf4", "Maf")), State)

write.csv(
  pseudobulk_qc,
  file.path("results", "tables", "Supplementary_Table_2_pseudobulk_coverage.csv"),
  row.names = FALSE
)
```

# 8. Pseudobulk raw-count sensitivity analysis

```{r pseudobulk-sensitivity}
comparisons <- combn(state_order, 2, simplify = FALSE)
pseudobulk_peak_tables <- list()

for (tf in names(tf_objects)) {
  current_obj <- subset(tf_objects[[tf]], subset = celltype != "Undefined")
  DefaultAssay(current_obj) <- "peaks_common"

  pb_counts <- AggregateExpression(
    current_obj,
    assays = "peaks_common",
    group.by = "celltype",
    slot = "counts",
    return.seurat = FALSE
  )$peaks_common

  cpm_matrix <- sweep(pb_counts, 2, colSums(pb_counts), FUN = "/") * 1e6

  for (comp in comparisons) {
    state_1 <- comp[1]
    state_2 <- comp[2]

    if (!all(c(state_1, state_2) %in% colnames(pb_counts))) next

    pseudobulk_peak_tables[[paste(tf, state_1, state_2, sep = "__")]] <- tibble(
      Peak = rownames(pb_counts),
      TF = tf,
      Comparison = paste0(state_1, "_vs_", state_2),
      Count_State1 = as.numeric(pb_counts[, state_1]),
      Count_State2 = as.numeric(pb_counts[, state_2]),
      CPM_State1 = as.numeric(cpm_matrix[, state_1]),
      CPM_State2 = as.numeric(cpm_matrix[, state_2]),
      Log2FC = log2(
        (as.numeric(cpm_matrix[, state_1]) + 1) /
          (as.numeric(cpm_matrix[, state_2]) + 1)
      )
    )
  }
}

pseudobulk_diff_table_full <- bind_rows(pseudobulk_peak_tables)
count_thresholds <- c(1, 2, 5, 10, 20)

threshold_sensitivity <- map_dfr(count_thresholds, function(min_count) {
  pseudobulk_diff_table_full %>%
    group_by(TF, Comparison) %>%
    summarise(
      Min_count = min_count,
      N_eligible = sum(
        Count_State1 >= min_count & Count_State2 >= min_count
      ),
      N_differential = sum(
        Count_State1 >= min_count & Count_State2 >= min_count &
          abs(Log2FC) > 1
      ),
      .groups = "drop"
    ) %>%
    mutate(
      Pct_differential = ifelse(
        N_eligible > 0,
        100 * N_differential / N_eligible,
        NA_real_
      )
    )
})

trajectory_comparisons <- c(
  "TFH_vs_TR1.1",
  "TR1.1_vs_TR1.2",
  "TR1.2_vs_Foxp3+TR1.2"
)

threshold_sensitivity_trajectory <- threshold_sensitivity %>%
  filter(Comparison %in% trajectory_comparisons)

write.csv(
  threshold_sensitivity_trajectory,
  file.path("results", "tables", "threshold_sensitivity_long.csv"),
  row.names = FALSE
)
```

# 9. Balanced resampling: BCL6 TFH vs TR1.1

This robustness analysis is intentionally restricted to the representative BCL6 TFH-versus-TR1.1 comparison used in the final thesis.

```{r balanced-resampling}
bcl6_balanced <- subset(
  tf_objects[["Bcl6"]],
  subset = celltype %in% c("TFH", "TR1.1") & nCount_peaks_common > 0
)
DefaultAssay(bcl6_balanced) <- "peaks_common"

counts_mat <- GetAssayData(
  bcl6_balanced,
  assay = "peaks_common",
  slot = "counts"
)

cells_tfh <- WhichCells(bcl6_balanced, expression = celltype == "TFH")
cells_tr11 <- WhichCells(bcl6_balanced, expression = celltype == "TR1.1")

detection_support <- tibble(
  Peak = rownames(counts_mat),
  N_detected_TFH = Matrix::rowSums(counts_mat[, cells_tfh, drop = FALSE] > 0),
  N_detected_TR11 = Matrix::rowSums(counts_mat[, cells_tr11, drop = FALSE] > 0)
) %>%
  mutate(
    Pct_detected_TFH = 100 * N_detected_TFH / length(cells_tfh),
    Pct_detected_TR11 = 100 * N_detected_TR11 / length(cells_tr11)
  )

run_balanced_resampling <- function(
    counts_mat,
    cells_state1,
    cells_state2,
    min_cell_support,
    sampling_fraction = 0.80,
    n_iterations = 100,
    seed = 123
) {
  eligible <- detection_support %>%
    filter(
      N_detected_TFH >= min_cell_support |
        N_detected_TR11 >= min_cell_support
    ) %>%
    pull(Peak)

  filtered_mat <- counts_mat[eligible, , drop = FALSE]
  n_sample <- floor(
    sampling_fraction * min(length(cells_state1), length(cells_state2))
  )

  set.seed(seed)
  iterations <- vector("list", n_iterations)

  for (i in seq_len(n_iterations)) {
    sampled_1 <- sample(cells_state1, n_sample, replace = FALSE)
    sampled_2 <- sample(cells_state2, n_sample, replace = FALSE)

    counts_1 <- Matrix::rowSums(filtered_mat[, sampled_1, drop = FALSE])
    counts_2 <- Matrix::rowSums(filtered_mat[, sampled_2, drop = FALSE])

    # Library sizes are calculated from the complete common peak feature space.
    library_1 <- sum(counts_mat[, sampled_1, drop = FALSE])
    library_2 <- sum(counts_mat[, sampled_2, drop = FALSE])

    cpm_1 <- counts_1 / library_1 * 1e6
    cpm_2 <- counts_2 / library_2 * 1e6

    iterations[[i]] <- tibble(
      Peak = names(counts_1),
      Iteration = i,
      Count_TFH = as.numeric(counts_1),
      Count_TR11 = as.numeric(counts_2),
      CPM_TFH = as.numeric(cpm_1),
      CPM_TR11 = as.numeric(cpm_2),
      # Positive values indicate higher occupancy in TR1.1.
      Log2FC = log2((cpm_2 + 1) / (cpm_1 + 1))
    )
  }

  long <- bind_rows(iterations)

  summary <- long %>%
    group_by(Peak) %>%
    summarise(
      Median_log2FC = median(Log2FC, na.rm = TRUE),
      Q1_log2FC = quantile(Log2FC, 0.25, na.rm = TRUE),
      Q3_log2FC = quantile(Log2FC, 0.75, na.rm = TRUE),
      Pct_iterations_FC_gt1 = 100 * mean(Log2FC > 1),
      Pct_iterations_FC_lt_minus1 = 100 * mean(Log2FC < -1),
      .groups = "drop"
    ) %>%
    left_join(detection_support, by = "Peak")

  candidates <- summary %>%
    filter(
      abs(Median_log2FC) > 1,
      Pct_iterations_FC_gt1 >= 80 |
        Pct_iterations_FC_lt_minus1 >= 80
    ) %>%
    mutate(
      Direction = case_when(
        Pct_iterations_FC_gt1 >= 80 ~ "Higher in TR1.1",
        Pct_iterations_FC_lt_minus1 >= 80 ~ "Higher in TFH"
      ),
      Max_detected_cells = pmax(N_detected_TFH, N_detected_TR11)
    )

  list(
    minimum_support = min_cell_support,
    eligible_peaks = eligible,
    long = long,
    summary = summary,
    candidates = candidates
  )
}

resampling_3 <- run_balanced_resampling(
  counts_mat,
  cells_tfh,
  cells_tr11,
  min_cell_support = 3
)

resampling_5 <- run_balanced_resampling(
  counts_mat,
  cells_tfh,
  cells_tr11,
  min_cell_support = 5
)

balanced_table_A <- bind_rows(
  tibble(
    `Minimum cell support` = ">=3 cells",
    `Eligible peaks` = length(resampling_3$eligible_peaks),
    `Reproducible candidates` = nrow(resampling_3$candidates),
    `Higher in TFH` = sum(resampling_3$candidates$Direction == "Higher in TFH"),
    `Higher in TR1.1` = sum(resampling_3$candidates$Direction == "Higher in TR1.1")
  ),
  tibble(
    `Minimum cell support` = ">=5 cells",
    `Eligible peaks` = length(resampling_5$eligible_peaks),
    `Reproducible candidates` = nrow(resampling_5$candidates),
    `Higher in TFH` = sum(resampling_5$candidates$Direction == "Higher in TFH"),
    `Higher in TR1.1` = sum(resampling_5$candidates$Direction == "Higher in TR1.1")
  )
)

balanced_table_B <- resampling_5$candidates %>%
  mutate(
    `Maximum cellular support` = case_when(
      Max_detected_cells >= 20 ~ ">=20 cells",
      Max_detected_cells >= 10 ~ "10-19 cells",
      Max_detected_cells >= 7 ~ "7-9 cells",
      TRUE ~ "5-6 cells"
    ),
    `Maximum cellular support` = factor(
      `Maximum cellular support`,
      levels = c("5-6 cells", "7-9 cells", "10-19 cells", ">=20 cells")
    )
  ) %>%
  count(`Maximum cellular support`, name = `Number of candidates`) %>%
  mutate(
    `Percentage of candidates` = round(
      100 * `Number of candidates` / sum(`Number of candidates`),
      1
    )
  )

write.csv(
  balanced_table_A,
  file.path("results", "tables", "Supplementary_Table_4A_balanced_resampling.csv"),
  row.names = FALSE
)
write.csv(
  balanced_table_B,
  file.path("results", "tables", "Supplementary_Table_4B_cellular_support.csv"),
  row.names = FALSE
)
```

# 10. Thesis figures

## Figure 4A: reference-mapped cellular states

```{r figure-4a}
umap_plots <- imap(tf_objects, function(obj, tf) {
  DimPlot(
    obj,
    group.by = "celltype",
    reduction = "umap_rna",
    cols = state_cols,
    pt.size = 0.8
  ) +
    ggtitle(toupper(tf)) +
    labs(x = "UMAP 1", y = "UMAP 2")
})

figure_4a <- wrap_plots(umap_plots, nrow = 1, guides = "collect") &
  theme(legend.position = "bottom")

figure_4a
```

## Figure 4B: state-specific module scores

```{r figure-4b}
score_features <- c(
  "TFH_Score",
  "TR1.1_Score",
  "TR1.2_Score",
  "Foxp3_TR1.2_Score"
)
score_titles <- c(
  "TFH signature",
  "TR1.1 signature",
  "TR1.2 signature",
  "Foxp3+TR1.2 signature"
)

feature_panels <- imap(tf_objects, function(obj, tf) {
  plots <- map2(score_features, score_titles, function(feature, title) {
    FeaturePlot(
      obj,
      features = feature,
      reduction = "umap_rna",
      cols = c("lightgrey", "red"),
      order = TRUE
    ) +
      ggtitle(title) +
      labs(x = "UMAP 1", y = "UMAP 2")
  })

  wrap_plots(plots, ncol = 2) +
    plot_annotation(title = toupper(tf))
})

figure_4b <- wrap_plots(feature_panels, nrow = 1)
figure_4b
```

## Figure 5: state-specific gene-expression dot plots

```{r figure-5}
signature_genes <- unique(unlist(gene_signatures, use.names = FALSE))

dotplots <- imap(tf_objects, function(obj, tf) {
  DefaultAssay(obj) <- "RNA"
  obj$celltype <- factor(obj$celltype, levels = state_order_with_undefined)

  DotPlot(
    obj,
    features = signature_genes,
    group.by = "celltype",
    cols = "RdBu"
  ) +
    coord_flip() +
    ggtitle(toupper(tf)) +
    labs(x = NULL, y = NULL) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
})

figure_5 <- wrap_plots(dotplots, nrow = 1, guides = "collect") +
  plot_annotation(tag_levels = "A")
figure_5
```

## Figure 6: consensus peak set and pseudobulk coverage

```{r figure-6}
tf_peak_colors <- c(
  "Bcl6" = "#E64B35",
  "Irf4" = "#4DBBD5",
  "Maf" = "#00A087",
  "Union (consensus)" = "#8491B4"
)

peak_width_df <- bind_rows(
  imap_dfr(original_peak_ranges, function(gr, tf) {
    tibble(TF = tf, Width = width(gr))
  }),
  tibble(
    TF = "Union (consensus)",
    Width = width(consensus_peaks_unfiltered)
  )
)

plot_a <- ggplot(peak_width_df, aes(Width, fill = TF, color = TF)) +
  geom_density(alpha = 0.3, linewidth = 0.7) +
  geom_vline(xintercept = 2000, linetype = "dashed") +
  scale_x_log10(labels = comma) +
  scale_fill_manual(values = tf_peak_colors) +
  scale_color_manual(values = tf_peak_colors) +
  labs(
    title = "Peak width distribution per transcription factor",
    x = "Peak width (bp, log10 scale)",
    y = "Density"
  ) +
  theme_classic()

cell_count_df <- pseudobulk_qc %>%
  transmute(TF, State, N_cells)

plot_b <- ggplot(cell_count_df, aes(State, N_cells, fill = State)) +
  geom_col(color = "black", linewidth = 0.3, width = 0.7) +
  geom_text(aes(label = comma(N_cells)), vjust = -0.4, size = 3) +
  facet_wrap(~TF, nrow = 1, scales = "free_y") +
  scale_fill_manual(values = state_cols) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Cell numbers per state", x = NULL, y = "Cells per pseudobulk") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

plot_c <- ggplot(
  pseudobulk_qc,
  aes(State, N_detected_peaks, fill = State)
) +
  geom_col(color = "black", linewidth = 0.3, width = 0.7) +
  geom_text(aes(label = comma(N_detected_peaks)), vjust = -0.4, size = 3) +
  facet_wrap(~TF, nrow = 1, scales = "free_y") +
  scale_fill_manual(values = state_cols) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Detected consensus peaks per cell-state pseudobulk",
    x = NULL,
    y = "Consensus peaks with detected CUT&Tag signal"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

figure_6 <- (plot_a / plot_b / plot_c) + plot_annotation(tag_levels = "A")
figure_6

ggsave(
  file.path("figures", "Figure_6_consensus_peakset_and_coverage.png"),
  figure_6,
  width = 10.5,
  height = 12,
  dpi = 300
)
```

# 11. Supplementary threshold-sensitivity figure and compact table

```{r supplementary-thresholds}
supp_fig_3a <- ggplot(
  threshold_sensitivity_trajectory,
  aes(Min_count, N_eligible, group = Comparison)
) +
  geom_line() +
  geom_point() +
  facet_wrap(~TF, scales = "free_y") +
  scale_x_continuous(breaks = count_thresholds) +
  labs(
    x = "Minimum pseudobulk counts required in both states",
    y = "Eligible peaks"
  ) +
  theme_classic()

supp_fig_3b <- ggplot(
  threshold_sensitivity_trajectory,
  aes(Min_count, Pct_differential, group = Comparison)
) +
  geom_line() +
  geom_point() +
  facet_wrap(~TF, scales = "free_y") +
  scale_x_continuous(breaks = count_thresholds) +
  labs(
    x = "Minimum pseudobulk counts required in both states",
    y = "Differential peaks among eligible peaks (%)"
  ) +
  theme_classic()

supp_figure_3 <- supp_fig_3a / supp_fig_3b + plot_annotation(tag_levels = "A")

ggsave(
  file.path("figures", "Supplementary_Figure_3_threshold_sensitivity.png"),
  supp_figure_3,
  width = 10,
  height = 8,
  dpi = 300
)

supp_table_3 <- threshold_sensitivity_trajectory %>%
  mutate(
    Threshold = ifelse(Min_count == 1, ">=1 count", paste0(">=", Min_count, " counts")),
    Result = ifelse(
      N_eligible > 0,
      paste0(N_eligible, " / ", N_differential, " (", round(Pct_differential, 1), "%)"),
      "0 / 0"
    )
  ) %>%
  select(TF, Comparison, Threshold, Result) %>%
  pivot_wider(names_from = Threshold, values_from = Result)

write.csv(
  supp_table_3,
  file.path("results", "tables", "Supplementary_Table_3_threshold_sensitivity.csv"),
  row.names = FALSE
)
```

# 12. Save final analysis objects and reproducibility information

```{r save-results}
saveRDS(
  tf_objects,
  file.path("results", "objects", "tf_objects_final.rds")
)
saveRDS(
  consensus_peaks_final,
  file.path("results", "objects", "consensus_peaks_final.rds")
)
saveRDS(
  resampling_3,
  file.path("results", "objects", "balanced_resampling_3cells.rds")
)
saveRDS(
  resampling_5,
  file.path("results", "objects", "balanced_resampling_5cells.rds")
)

writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```
